---
title: "Diversité_bac"
author: "Divya"
date: "2023-01-30"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
```


### Load libraries
```{r message=TRUE}
library(vegan)
library(tidyr)
library(dplyr)
library(FSA)
library(ggplot2)
```


### Loading all files needed here and seperating certain dataframes
```{r}
# Load ASV table and metadata
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)

# remove the mysterious "is.neg" column
metadata <- metadata %>% select(-is.neg)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"
```

### Calculate the ALPHA DIVERSITY
```{r}
# Calculate the alpha diversity and put it in the metadata data fame
metadata$alpha_diversity <- diversity(ASV_table, index = "shannon")
metadata$richness <- specnumber(ASV_table)
```

### Separate the Filters from the Rocks. We'll do dataframes WITHOUT the Terrain data 
```{r}
# So these will be with the Terrain (in both dataframe)
Filters_all <- metadata[(metadata$SampleType == "Filtre"), ]
Filters_all$Nom <- gsub("F-", "", Filters_all$Nom)

Rocks_all <- rbind(subset(metadata, SampleType == "Roche"),
                   subset(metadata, Bioreacteur == "Terrain"))
Rocks_all$Nom <- gsub("R-", "", Rocks_all$Nom)


# And these will be the ones without the Terrain 
Filters_noT <- Filters_all[!(Filters_all$Bioreacteur == "Terrain"), ]
Rocks_noT <- subset(metadata, SampleType == "Roche")

# And let's make a dataframe with Filters and Rocks, but no Terrain
RFonly <- metadata[!(metadata$Bioreacteur == "Terrain"), ]

# To merge our geochemical data to the reste of our metadata, I'll have to make another dataframe
Analyse <- read.table("/home/dpatel/Analyse_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)
Filtres_noT_geo <- merge(Filters_noT, Analyse, by = c("Bioreacteur", "Jour"), sort = TRUE)
```


### Visualizing our alpha diversity WITHOUT our Terrain
```{r}
# Change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
RFonly$Bioreacteur <- ifelse(grepl("^B", RFonly$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", RFonly$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", RFonly$Bioreacteur), "Bioreactor 3", "Field")))

# Here, we want our bioreactors to have a similar diversity since these are supposed to be triplicates. 
# We'll start off with boxplots for our Filters and Rocks WITHOUT our Terrain data
RF_box <- ggplot(RFonly, aes(x = Bioreacteur, y = alpha_diversity, fill = SampleType)) + 
  geom_boxplot() +
  labs(x = "Bioreactors", y = "Shannon index for Bacteria") +
  scale_fill_manual(labels = c("Planktonic", "Sessile"), values = c("#2A9D8F", "#F4A261"), name = "Community") +
 theme(text = element_text(size = 32), legend.text = element_text(size = 30))
RF_box
ggsave("/home/dpatel/sequences/Bac/bac_Figures/Alpha_bac.tiff", width = 12, height = 9, plot = last_plot())


# Looks like our Rocks have a higher diversity than our Filters.
# Between Bioreactors,whether we look at our Filters or our Rocks, they seem pretty similar.
```

### Looking at the normalicy of our data
```{r}
# Making a plot with our alpha diversities data for our Filters and our Rocks
hist(Filters_noT$alpha_diversity)
qqnorm(Filters_noT$alpha_diversity)
qqline(Filters_noT$alpha_diversity, distribution = qnorm)
shapiro.test(Filters_noT$alpha_diversity)
### W = 0.9882, p-value = 0.9787

hist(Rocks_noT$alpha_diversity)
shapiro.test(Rocks_noT$alpha_diversity)
### W = 0.95556, p-value = 0.2376
### In both cases, the p-value is greater than 0.05. So the data follows a normal distribution.

# Let's look at the RFonly
shapiro.test(RFonly$alpha_diversity)
### W = 0.91465, p-value = 0.0004705
### So it doesn't follow a normal distribution
```

### Checking the Homogeneity of Variance
```{r}
bartlett.test(alpha_diversity ~ Bioreacteur, data = Filters_noT)
### Bartlett's K-squared = 1.736, df = 2, p-value = 0.4198
### They have the same variance among Bioreactors

bartlett.test(alpha_diversity ~ Bioreacteur, data = Rocks_noT)
### Bartlett's K-squared = 7.3023, df = 2, p-value = 0.02596
### The variance is different among Bioreactors

fligner.test(alpha_diversity ~ SampleType, data = RFonly)
### Fligner-Killeen:med chi-squared = 12.595, df = 1, p-value = 0.0003869
### The variance is different among SampleType
```


### Comparing diversity between Bioreactors (No Terrain)
```{r}
# To compare our Bioreactors for our Rock and Filter, we'll do a Krulskall Wallis test
kruskal.test(alpha_diversity ~ Bioreacteur, data = Filters_noT)
### Kruskal-Wallis chi-squared = 4.2606, df = 2, p-value = 0.1188
### There is no significant difference between the Bioreactors

kruskal.test(alpha_diversity ~ Bioreacteur, data = Rocks_noT)
### Kruskal-Wallis chi-squared = 10.606, df = 2, p-value = 0.004978
### There is a significant difference between the Bioreactors

# To compare our SampleType, we'll do a Krulskall Wallis test
kruskal.test(alpha_diversity ~ SampleType, data = RFonly)
### Kruskal-Wallis chi-squared = 31.729, df = 1, p-value = 1.772e-08
### There is a significant difference between the SampleType
```


### Regression between our environmental factors and our alpha diversity
```{r}
# For this part, we're just using our Filters data without our Terrain in it. I'll have to make another dataframe since I'm missing some data.
Analyse <- read.table("/home/dpatel/Analyse_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)

Filtres_noT_geo <- merge(Filters_noT, Analyse, by = c("Bioreacteur", "Jour"), sort = TRUE)

# We'll do multiple models with the variables
geo_glm1 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 + Moy_NO2 + Moy_NOx + pH, data = Filtres_noT_geo)
geo_glm2 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 + Moy_NO2 + Moy_NOx, data = Filtres_noT_geo)
geo_glm3 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 + Moy_NO2 , data = Filtres_noT_geo)
geo_glm4 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 , data = Filtres_noT_geo)
geo_glm5 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 , data = Filtres_noT_geo)
geo_glm6 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC , data = Filtres_noT_geo)
geo_glm7 <- glm(alpha_diversity ~ Moy_DIC , data = Filtres_noT_geo)

# Checking which models is the best
AIC(geo_glm1,geo_glm2,geo_glm3,geo_glm4,geo_glm5,geo_glm6,geo_glm7)

# Results
#           df     AIC
# geo_glm1	9	  36.45443		
# geo_glm2	8	  34.45444		
# geo_glm3	7	  33.78771		
# geo_glm4	6	  33.10867		
# geo_glm5	5	  31.10966		
# geo_glm6	4	  33.75135		
# geo_glm7	3	  33.12171	


# The best model seems to be our model 5 (alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4).
# Let's compare the model to the other ones.
anova(geo_glm1, geo_glm5, test="Chisq")
anova(geo_glm2, geo_glm5, test="Chisq")
anova(geo_glm3, geo_glm5, test="Chisq")
anova(geo_glm6, geo_glm5, test="Chisq")
anova(geo_glm7, geo_glm5, test="Chisq")
# The geo_glm4 is not statistically different from the model 1, 2, 3 and 7, but is statistically different from the model 6 (but not by much).
# So it seems like the NH4 is an important parameter and that we must keep it.

# Now, let's look at the results of our glm and also we should look at the Residuals
summary(geo_glm5)
plot(geo_glm5)
```

