---
title: "Diversité_arc"
author: "Divya"
date: "2023-02-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
```


### Load libraries
```{r message=TRUE}
library(vegan)
library(tidyr)
library(dplyr)
library(FSA)
library(ggplot2)
```


### Calculate alpha diversity
```{r}
# Load ASV table and metadata
ASV_table <- t(read.table("/home/dpatel/sequences/Arc/arc_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("/home/dpatel/sequences/Arc/arc_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
# remove the mysterious "is.neg" column
metadata <- metadata %>% select(-is.neg)

# Change the name of the first column to "Nom"
colnames(metadata)[1] <- "Nom"

# Calculate the alpha diversity and put it in the metadata data fame
metadata$alpha_diversity <- diversity(ASV_table, index = "shannon")
metadata$richness <- specnumber(ASV_table)

```


### Separate the Filters from the Rocks. We'll do dataframes WITH and WITHOUT the Terrain (field) data 
```{r}
# So these will be with the Terrain (in both dataframe)
Filters_all <- metadata[(metadata$SampleType == "Filtre"), ]
Filters_all$Nom <- gsub("F-", "", Filters_all$Nom)

# And these will be the ones without the Terrain 
Filters_noT <- Filters_all[!(Filters_all$Bioreacteur == "Terrain"), ]
Filters_noT_1 <- Filters_all[!(Filters_all$Bioreacteur == "Terrain"), ]
```


### Visualizing our alpha diversity WITHOUT our Terrain
```{r}
# Change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
Filters_noT$Bioreacteur <- ifelse(grepl("^B", Filters_noT$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", Filters_noT$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", Filters_noT$Bioreacteur), "Bioreactor 3", "Field")))

# Here, we want our bioreactors to have a similar diversity since these are supposed to be triplicates. 
# We'll start off with boxplots for our Filters and Rocks WITHOUT our Terrain data
RF_box <- ggplot(Filters_noT, aes(x = Bioreacteur, y = alpha_diversity, fill = SampleType)) + 
  geom_boxplot() +
  labs(x = "Bioreactor", y = "Shannon index for Archaea") +
  scale_fill_manual(labels = c("Planctonique"), values = c("#2A9D8F"), name = "Community") +
theme(text = element_text(size = 22), legend.text = element_text(size = 18))
RF_box
#ggsave("/home/dpatel/sequences/Arc/arc_Figures/Alpha.jpeg", width = 11, height = 9, plot = last_plot())

# Looks like our Rocks have a higher diversity than our Filters.
# Between Bioreactors,whether we look at our Filters or our Rocks, they seem pretty similar.
```

### Looking at the normalicy of our data
```{r}
# Making a plot with our alpha diversities data for our Filters and our Rocks
hist(Filters_noT$alpha_diversity)
qqnorm(Filters_noT$alpha_diversity)
qqline(Filters_noT$alpha_diversity, distribution = qnorm)
shapiro.test(Filters_noT$alpha_diversity)
### W = 0.80208, p-value = 0.001225

### The p-value is smaller than 0.05. So the data doesn't follows a normal distribution.
```

### Checking the Homogeneity of Variance
```{r}
fligner.test(alpha_diversity ~ Bioreacteur, data = Filters_noT)
### Fligner-Killeen:med chi-squared = 0.79133, df = 2, p-value = 0.6732
### The variance is the same among Bioreacteurs
```


### Comparing diversity between Bioreactors (No field data)
```{r}
# To compare our Bioreactors for our Rock and Filter, we'll do a Krulskall Wallis test
kruskal.test(alpha_diversity ~ Bioreacteur, data = Filters_noT)
### Kruskal-Wallis chi-squared = 0.38921, df = 2, p-value = 0.8232
### There is no significant difference between the Bioreactors

```

### Regression between our environmental factors and our alpha diversity
```{r}
# For this part, we're just using our Filters data without our Terrain in it. I'll have to make another dataframe since I'm missing some data.
Analyse <- read.table("/home/dpatel/Analyse_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)
Filtres_noT_geo <- merge(Filters_noT_1, Analyse, by = c("Bioreacteur", "Jour"), sort = TRUE)

# We'll do multiple models with the variables
geo_glm1 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 + Moy_NO2 + Moy_NOx, data = Filtres_noT_geo)
geo_glm2 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 + Moy_NO2 , data = Filtres_noT_geo)
geo_glm3 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 + Moy_NO3 , data = Filtres_noT_geo)
geo_glm4 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4 , data = Filtres_noT_geo)
geo_glm5 <- glm(alpha_diversity ~ Moy_DIC + Moy_DOC , data = Filtres_noT_geo)
geo_glm6 <- glm(alpha_diversity ~ Moy_DIC , data = Filtres_noT_geo)

# Checking which models is the best
AIC(geo_glm1,geo_glm2,geo_glm3,geo_glm4,geo_glm5,geo_glm6)

# Results
#           df     AIC
# geo_glm1	8	 -38.56878		
# geo_glm2	7	 -39.69096		
# geo_glm3	6	 -41.26544		
# geo_glm4	5	 -42.70575		
# geo_glm5	4	 -37.72118		
# geo_glm6	3	 -20.06878	


# The best model seems to be our model 4 (alpha_diversity ~ Moy_DIC + Moy_DOC + Moy_NH4). Let's compare the model to the other ones.
anova(geo_glm1, geo_glm4, test="Chisq")
anova(geo_glm2, geo_glm4, test="Chisq")
anova(geo_glm3, geo_glm4, test="Chisq")
anova(geo_glm5, geo_glm4, test="Chisq")
anova(geo_glm6, geo_glm4, test="Chisq")

# The geo_glm4 is not statistically different from the model 1, 2 and 3, but is statistically different from the model 5 and 6.
# So it seems like the NH4 and DOC are important parameters and that we must keep them.

# Now, let's look at the results of our glm and also we should look at the Residuals

summary(geo_glm4)
plot(geo_glm4)

```



