---
title: "Diversité bêta"
author: "Divya"
date: "2023-05-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
```


### Load libraries
```{r message=TRUE}
library(vegan)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stringr)
```


### doing a dissimilarity matrix (Bray-Kurtis)
```{r}
# Load ASV table and metadata
ASV_table <- t(read.table("/home/dpatel/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))

# Let's start by doing a dissimilarity matrix (Bray-Kurtis)
beta_dist <- vegdist(ASV_table, method = "bray")
write.csv(as.matrix(beta_dist), "/home/dpatel/sequences/Bac/bac_csv/dist_matrix.csv")
beta_dist <- read.table("/home/dpatel/sequences/Bac/bac_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = NULL,)
```


### Making a PCoA
```{r}
# Load the dissimilarity matrix
beta_dist <- read.table("/home/dpatel/sequences/Bac/bac_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = 1)

# Perform PCoA analysis using the "cmdscale" function
PCOA <- cmdscale(beta_dist, eig=TRUE, add=TRUE)

# Extract the coordinate results from the PCOA
pcoa_df <- data.frame(PCoA1 = PCOA$points[,1], PCoA2 = PCOA$points[,2], Sample = row.names(beta_dist))

pcoa_df$SampleType <- ifelse(grepl("^F", pcoa_df$Sample), "Filter",
                        ifelse(grepl("^R", pcoa_df$Sample), "Rock",
                               ifelse(grepl("^T|^P", pcoa_df$Sample), "Field", "Other")))

# Remove "R." and "F." from the sample names in the PCoA data frame and change the J of Jour to D (day)
pcoa_df$Sample <- gsub("^R\\.|^F\\.", "", pcoa_df$Sample)
pcoa_df$Sample <- gsub(".J", "-D", pcoa_df$Sample)

# Modify the Sample column, keeping "P1", "P2", "P3", and "P4" unchanged
pcoa_df$Sample <- ifelse(pcoa_df$Sample %in% c("P1", "P2", "P3", "P4"), pcoa_df$Sample, 
                        str_replace(pcoa_df$Sample, "(\\d+)$", function(match) as.character(2 * as.numeric(match))))

# Create a new column in the PCoA data frame that indicates the shape of each sample
pcoa_df$Shape <- ifelse(grepl("^B", pcoa_df$Sample), "circle",
                        ifelse(grepl("^J", pcoa_df$Sample), "diamond",
                         ifelse(grepl("^R", pcoa_df$Sample), "square", "triangle")))
pcoa_df$Sample<- gsub("Tourb", "Peat", pcoa_df$Sample)

percent_explained=100*PCOA$eig/sum(PCOA$eig)
reduced_percent=format(round(percent_explained[1:2], digits=2), nsmall=1, trim=TRUE) # reduce number of digits

# Visualize the PCoA plot using ggplot
p <- ggplot(pcoa_df, aes(x = PCoA1, y = PCoA2, color = SampleType, shape = Shape, label = Sample)) +
  geom_point(size = 3.5) +
  xlab(paste0("PCoA 1 (", reduced_percent[1], "% of variation)")) +
  ylab(paste0("PCoA 2 (", reduced_percent[2], "% of variation)")) +
  geom_text(hjust = 0, vjust = 0) +
  theme(legend.position = "right") +  # move legend to the right
  scale_color_manual(values = c("#116530","#21B6A8","#A45C40"), # Set colors
                     name = "Sample Type") +  # define legend title and labels
  scale_shape_manual(values = c(16, 17, 15, 18),  # Set the shape scheme
                      labels = c("Bioreactor 1", "Bioreactor 2", "Bioreactor 3", "Field"), # Set the labels
                     name = "Bioreactor") 
  
p + xlim(-0.5, 0.8) + # Set x-axis limits
theme(text = element_text(size = 22), legend.text = element_text(size = 18))

ggsave("/home/dpatel/sequences/Bac/bac_Figures/PCOA_beta4.tiff", width = 8, height = 6, plot = last_plot())
```



###### Correlation between our environmental factors and our beta diversity (PERMANOVA) ######
### NO TERRAIN DATA
### Loading all files needed here
```{r}
# Load ASV table and metadatas
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
```


### Doing some cleaning in the dataframes
```{r}
# Remove the mysterious "is.neg" column and remove the unecessary column in metadata 
metadata <- metadata %>% select(-is.neg, -Rugosite1, -Rugosite2, -concADN)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"
```


### PERMANOVA FOR FILTERS ###

# After taking a look at the PCoA, we'll separate the data based on the Jours. 
# So the days will be J1-J4, J5-J6 and J7-J12
```{r}
rm(list=ls())

# Load ASV table and metadatas
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
beta_dist <- read.table("/home/dpatel/sequences/Bac/bac_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = NULL)
colnames(beta_dist) <- c("Nom", beta_dist$X)
```

### Doing some cleaning in the dataframes
```{r}
# Remove the mysterious "is.neg" column and remove the unecessary column in metadata 
metadata <- metadata %>% select(-is.neg, -Rugosite1, -Rugosite2, -concADN)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"
```

### Doing the PERMANOVA (SEPARATE BASED ON JOURs)
```{r}
set.seed(123)
# Separate the Filtres based on the Jours 
meta14  <- metadata[metadata$Jour %in% c("2", "4", "6", "8"), ]
meta56  <- metadata[metadata$Jour %in% c("10", "12"), ]
meta711 <- metadata[metadata$Jour %in% c("14", "16", "18", "20","22"), ]

# Now we have to join the metadata dataframes with the matrix and keep the lower half of the dist
meta14_dist <- inner_join(meta14, beta_dist, by = "Nom")
dist14 <- meta14_dist %>% select(all_of(meta14_dist$Nom)) %>% as.dist()

meta56_dist <- inner_join(meta56, beta_dist, by = "Nom")
dist56 <- meta56_dist %>% select(all_of(meta56_dist$Nom)) %>% as.dist()

meta711_dist <- inner_join(meta711, beta_dist, by = "Nom")
dist711 <- meta711_dist %>% select(all_of(meta711_dist$Nom)) %>% as.dist()


# Run the adonis2 function and look at the results
Corr14 <- adonis2(dist14 ~ Bioreacteur + Jour, data = meta14_dist, by="margin")
print(Corr14)
#              Df SumOfSqs     R2      F     Pr(>F)    
# Bioreacteur  2  0.10429   0.19131  1.6162  0.082 .  
# Jour         1  0.18274   0.33522  5.6641  0.001 ***
# Residual     8  0.25810   0.47347                  
# Total       11  0.54513   1.00000    

Corr56 <- adonis2(dist56 ~ Bioreacteur + Jour, data = meta56_dist, by="margin")
print(Corr56)
#              Df  SumOfSqs     R2       F     Pr(>F)  
# Bioreacteur  2  0.108084   0.52754  3.4987  0.01667 *
# Jour         1  0.065905   0.32167  4.2666  0.01667 *
# Residual     2  0.030893   0.15078                 
# Total        5  0.204882   1.00000  

Corr711 <- adonis2(dist711 ~ Bioreacteur + Jour, data = meta711_dist, by="margin")
print(Corr711)
#             Df  SumOfSqs     R2       F    Pr(>F)    
# Bioreacteur  2  0.29413   0.32333  3.2372  0.002 ** 
# Jour         1  0.24824   0.27288  5.4642  0.001 ***
# Residual     8  0.36344   0.39951                  
# Total       11  0.90970   1.00000   
```

### Doing the PERMANOVA (ALL FILTERS COMBINED)
```{r}
# Keeping the Filtres only
meta_filter <- metadata[(metadata$SampleType == "Filtre"), ]
meta_filter <- meta_filter[!(meta_filter$Bioreacteur == "Terrain"), ]

# Now we have to join the metadata dataframes with the matrix and keep the lower half of the dist
meta_dist <- inner_join(meta_filter, beta_dist, by = "Nom")
dist <- meta_dist %>% select(all_of(meta_dist$Nom)) %>% as.dist()

# Run the adonis2 function and look at the results
Corr <- adonis2(dist ~ Bioreacteur + Jour, data = meta_dist, by="margin")
print(Corr)
#             Df  SumOfSqs     R2       F     Pr(>F)    
# Bioreacteur  2   0.3631   0.08660   3.5476  0.009 ** 
# Jour         1   2.4944   0.59482  48.7365  0.001 ***
# Residual    26   1.3307   0.31732                   
# Total       29   4.1936   1.00000    
```


### PERMANOVA FOR LIFESTYLE ###
```{r}
meta_dist <- inner_join(metadata, beta_dist, by = "Nom")
dist <- meta_dist %>% select(all_of(meta_dist$Nom)) %>% as.dist()
Corr <- adonis2(dist ~ SampleType, data = meta_dist, by="margin")
print(Corr)
```


### PERMANOVA FOR ROCKS ###

# Here, we're going to add the mineralogy with the metadata
# Load ASV table and metadatas
```{r}
rm(list=ls())
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
beta_dist <- read.table("/home/dpatel/sequences/Bac/bac_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = NULL)
colnames(beta_dist) <- c("Nom", beta_dist$X)
```


### Doing some cleaning in the dataframes
```{r}
# Remove the mysterious "is.neg" column and remove the unecessary column in metadata 
metadata <- metadata %>% select(-is.neg, -concADN, -Jour)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"
```

### Doing the PERMANOVA
```{r}
# Remove the 'Terrain' data 
metadata <- metadata[metadata$SampleType == "Roche", ]

# Now we have to join the metadata dataframe with the matrix and keep the lower half of the dist
meta_dist <- inner_join(metadata, beta_dist, by = "Nom")
all_dist <- meta_dist %>% select(all_of(meta_dist$Nom)) %>% as.dist()

# Run the adonis2 function and look at the results
Corr <- adonis2(all_dist ~ Bioreacteur, data = meta_dist, by="margin")
print(Corr)
#             Df  SumOfSqs    R2      F     Pr(>F)    
#Bioreacteur  2   1.0687   0.45537 11.287  0.001 ***
#Residual    27   1.2782   0.54463                  
#Total       29   2.3469   1.00000                 
```





###### Doing a dbRDA ########
# For the dbRDA, The ASV dataframe should have samples as rows and ASVs as columns
# As for the metadata dataframe, it should have samples as rows and metadata variables as columns.


### FILTERS ONLY
### Loading dataframes and doing some cleaning
```{r}
rm(list=ls())

# Load the ASV table and the metadata
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)

# Remove the mysterious "is.neg" column and remove the Rugosite1 and Rugosite2 because we don't need them
metadata <- metadata %>% select(-is.neg,-Rugosite1,-Rugosite2)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"

# Taking all my Filter data from metadata
Meta_filters_all <- metadata[(metadata$SampleType == "Filtre"), ]

# Merging the geochemical data to the reste of our metadata.
Analyse <- read.table("/home/dpatel/Analyse_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)
Meta_filters_all <- merge(Meta_filters_all, Analyse, by = c("Bioreacteur", "Jour"), sort = TRUE)

# Change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
Meta_filters_all$Bioreacteur <- ifelse(grepl("^B", Meta_filters_all$Bioreacteur), "Bio1",
                        ifelse(grepl("^J", Meta_filters_all$Bioreacteur), "Bio2",
                         ifelse(grepl("^R", Meta_filters_all$Bioreacteur), "Bio3", "Field")))

# Taking the ASV table and keeping the same samples as in the Filtres_noT_geo metadata.
# We'll also transform the raw ASV data into relative abundance 
ASV_table <- ASV_table[Meta_filters_all$Nom, ]
ASV_table <- decostand(ASV_table, method = "hellinger")

# Removing "F." from the sample names for the ASV table and the Meta_filters_all dataframe 
rownames(ASV_table) <- gsub("F\\.", "", rownames(ASV_table))
Meta_filters_all$Nom <- gsub("F\\.", "", Meta_filters_all$Nom)

# Change the J of Jour to D (day)
Meta_filters_all$Nom <- gsub(".J", "-D", Meta_filters_all$Nom)

# Extract numbers from the end of Sample names and multiply by 2
Meta_filters_all <- Meta_filters_all %>%
mutate(Nom = str_replace(Nom, "(\\d+)$", function(match) as.character(2 * as.numeric(match))))

```

### Doing the Distance matrix and dbRDA 
```{r}

beta_dist <- vegdist(ASV_table, method = "bray")

dbRDA_result <- capscale(beta_dist ~ Moy_DIC + Moy_DOC + Moy_NO2 + Moy_NOx + Moy_NH4 + Moy_NO3 + pH , data = Meta_filters_all)
```

### Looking at the plot
```{r}
# Doing a simple plot
plot(dbRDA_result)
```

# Doing some statistical test
```{r}
set.seed(123)
anova(dbRDA_result) 

# test axes for significance
anova(dbRDA_result, by = "axis", perm.max = 1000) 

# test for significant variables from the matadata
anova(dbRDA_result, by = "terms", permu = 1000) 

summary(dbRDA_result)
```

### Ok, Let's see if we can simplify our dbRDA
```{r}
#So step() is going to find the simplest models we can have: taking away variables that aren't too important
mod1 <- step(dbRDA_result, scope = formula(dbRDA_result), criteria = "AIC", test = "perm", direction = "both") 
summary(mod1)

mod2 <- capscale(beta_dist ~ Moy_DIC + Moy_NH4, data = Meta_filters_all)
```

### Doing statistical analysis on the simpler model
```{r}
# Set a random seed for reproducibility
set.seed(123) 

# Looking for some collinearity
vif.cca(mod2)
### Generally, VIF values greater than 5 or 10 are considered indicative of significant collinearity.
### Ok, so we don't see any "inflation" in the numbers so there isn't any collinearity

# test axes for significance
anova(mod2, by = "axis", perm.max = 1000) 

# test for significant variables from the metadata
anova(mod2, by = "terms", permu = 1000) 
#         Df SumOfSqs      F Pr(>F)   
#Moy_DIC   1  0.41656 3.8056  0.003 **
#Moy_NH4   1  0.20453 1.8685  0.056 . 
#Residual 21  2.29864 

#summary(mod2)

#plot(mod2)
```


### Plotting
```{r}

# Extract the site scores from dbRDA result
site_scores <- scores(mod2, display = "sites")
site_scores_df <- as.data.frame(site_scores)

# Extract the variable scores from dbRDA result
variable_scores <- scores(mod2, display = "bp")
variable_scores_df <- as.data.frame(variable_scores)

# Get the dimension names from the site scores
dim_names <- colnames(site_scores_df)

# Calculate the eigenvalues of the dbRDA result
eigenvals <- eigenvals(mod2)

# Calculate the percentage of variance explained by each dimension
var_explained <- eigenvals / sum(eigenvals) * 100

# Create a dataframe combining site scores and metadata
plot_data <- cbind.data.frame(site_scores_df, Meta_filters_all)

dbRDAfilter <-ggplot() +
  geom_point(data = plot_data, aes(x = CAP1, y = CAP2, color = Bioreacteur), size = 2) + #adding the points of our sample
  scale_color_manual(values = c("#21B6A8","#A45C40","#116530"), name = "Bioreactor",) + 
  
  geom_segment(data = variable_scores_df, aes(x = 0, y = 0, xend = CAP1, yend = CAP2), 
               arrow = arrow(length = unit(0.15, "cm")), color = "black") + #adding the arrows for the environmental variables
  xlab(paste(dim_names[1], " (", round(var_explained[1], 2), "%)")) + #naming the x axis
  ylab(paste(dim_names[2], " (", round(var_explained[2], 2), "%)")) + #naming the y axis
  geom_hline(yintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_vline(xintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_text(data = plot_data, aes(x = CAP1, y = CAP2, label = Nom), hjust = 0.5, vjust = -0.5, size = 3.5) + #adding the sample name
  theme_minimal() +
  theme(text = element_text(size = 15)) 

dbRDAfilter
ggsave("/home/dpatel/sequences/Bac/bac_Figures/Bac_dbRDA_filter.jpeg", width = 6, height = 4, plot = last_plot())
```



### ROCKS ONLY

### Loading dataframes and doing some cleaning

### Loading all files needed here
```{r}
rm(list=ls())

# Load ASV table and metadatas
ASV_table <- t(read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
mineral <-read.table("~/Mineralogie_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)
```


### Doing some cleaning in the dataframes
```{r}

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"

# Since our mineralogy is in percentage, we'll divide the values by 100.
mineral <- mineral %>% mutate_at(vars(X.IronOxide, X.Quartz, X.BlackQuartz, X.RedFeldpath, X.WhiteFeldpath, X.Mica, X.Undetermined), ~ . / 100)

# Also, since BlackQuartz and Quartz have the same molecular composition, we're going to add them into the Quartz column.
mineral <- mineral %>% mutate(X.Quartz = X.Quartz + X.BlackQuartz)
```

### Merging the metadata and mineral dataframes
```{r}
# Merging metadata with mineral dataframe so everything is together
# We'll then just keep the Roche samples
meta_mineral <- merge(metadata, mineral, by = "Nom", all.x = TRUE)

meta_mineral <- meta_mineral[(meta_mineral$SampleType == "Roche"), ]

# Remove the mysterious "is.neg" column and remove the Rugosite1 and Rugosite2 in Metadata since they will be in the Mineral dataframe
meta_mineral <- meta_mineral %>% select(-is.neg, -Rugosite1, -Rugosite2, -X.BlackQuartz, -Total, -Jour, -SampleType, -concADN)

# Change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
meta_mineral$Bioreacteur <- ifelse(grepl("^B", meta_mineral$Bioreacteur), "Bio1",
                        ifelse(grepl("^J", meta_mineral$Bioreacteur), "Bio2",
                         ifelse(grepl("^R", meta_mineral$Bioreacteur), "Bio3", "")))
```

# Taking the ASV table and keeping the same samples as in the meta_mineral metadata
```{r}
ASV_table <- ASV_table[meta_mineral$Nom, ]

# Removing "R." from the sample names for the meta_mineral dataframe 
meta_mineral$Nom <- gsub("R\\.", "", meta_mineral$Nom)

# We'll also transform the raw ASV data into relative abundance 
ASV_table <- decostand(ASV_table, method = "hellinger")
```


### Doing the Distance matrix and dbRDA 
```{r}
beta_dist <- vegdist(ASV_table, method = "bray")

dbRDA_result <- capscale(beta_dist ~ X.IronOxide + X.Quartz + X.RedFeldpath + X.WhiteFeldpath + X.Mica + X.Undetermined + Roughness, data = meta_mineral)
```

### Looking at the plot
```{r}
# Doing a simple plot
plot(dbRDA_result)
```

# Doing some statistical test
```{r}
set.seed(123)

anova(dbRDA_result) 

# test axes for significance
anova(dbRDA_result, by="axis", perm.max=1000) 

# test for significant variables from the matadata
anova(dbRDA_result, by="terms", permu=1000) 

summary(dbRDA_result)

```

### Ok, Let's see if we can simplify our dbRDA
```{r}
set.seed(123)
#So step() is going to find the simplest models we can have
mod1 <- step(dbRDA_result, scope = formula(dbRDA_result), test = "perm") 
summary(mod1)

# Looking for some collinearity
vif.cca(mod1)
###Ok, so we see an "inflation" in the numbers so there is collinearity

mod2 <- capscale(beta_dist ~ X.Quartz + X.RedFeldpath + X.Mica + Roughness, data = meta_mineral)
#summary(mod2)
vif.cca(mod2)
anova(mod2, by="terms", permu=1000)
#plot(mod1)
plot(mod2)
#plot(dbRDA_result)
```


### Doing statistical analysis on the simpler model
```{r}
predictor_vars <- scores(mod2, display = "sites")
cor_matrix <- cor(predictor_vars)
corrplot::corrplot(cor_matrix, method = "circle")
vif_values <- vegan::vif.cca(mod2)

```

### Plotting
```{r}
# Extract the site scores from dbRDA result
site_scores <- scores(mod2, display = "sites")
site_scores_df <- as.data.frame(site_scores)

# Extract the variable scores from dbRDA result
variable_scores <- scores(mod2, display = "bp")
variable_scores_df <- as.data.frame(variable_scores)

# Get the dimension names from the site scores
dim_names <- colnames(site_scores_df)

# Calculate the eigenvalues of the dbRDA result
eigenvals <- eigenvals(mod2)

# Calculate the percentage of variance explained by each dimension
var_explained <- eigenvals / sum(eigenvals) * 100

# Create a dataframe combining site scores and metadata
plot_data <- cbind.data.frame(site_scores_df, meta_mineral)

dbrdarock <- ggplot() +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_vline(xintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_point(data = plot_data, aes(x = CAP1, y = CAP2, color = Bioreacteur), size = 2) + #adding the points of our sample
  geom_segment(data = variable_scores_df, aes(x = 0, y = 0, xend = CAP1, yend = CAP2), 
               arrow = arrow(length = unit(0.15, "cm")), color = "black") + #adding the arrows for the environmental variables
  scale_color_manual(values = c("#21B6A8","#A45C40","#116530"), name = "Bioreactor",) + 
  xlab(paste(dim_names[1], " (", round(var_explained[1], 2), "%)")) + #naming the x axis
  ylab(paste(dim_names[2], " (", round(var_explained[2], 2), "%)")) + #naming the y axis
  geom_text(data = plot_data, aes(x = CAP1, y = CAP2, label = Nom), hjust = 0.5, vjust = -0.5, size = 3.5) + #adding the sample name
  theme_minimal() +
  theme(text = element_text(size = 15)) 

dbrdarock
ggsave("/home/dpatel/sequences/Bac/bac_Figures/Bac_dbRDA_rock.jpeg", width = 6, height = 4, plot = last_plot())
```



