---
title: "DADA2-Euc"
author: "Divya"
date: "2022-11-03"
output: html_document
---
## On va commencer par tout supprimer de notre data avec rm()

couper rarefaction à 1000

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, include = FALSE)

rm(list=ls())
```

## Opening DADA2 and setting our working directory for our Bacteria files

```{r message=FALSE, warning=FALSE, include=FALSE}

library(dada2); packageVersion("dada2")


#Setting a path to the FASTQ files
#Here I'm focusing on my Eukaryotes

path <-"~/sequences/Euc/euc_fastq"


#Just going to make sure the path is leading to the .fastq files

list.files(path)

#Alrigth, looks like it worked! ^.^

```

## Seperate the Forward and Reverse fastq files

```{r pressure, include=FALSE}
# Identify forward and reverse reads
# Filenames containing R1 are forward reads, R2 are reverse reads.

fnFs <- sort(list.files(path, pattern="R1_001.fastq", full.names = TRUE))
fnRs <- sort(list.files(path, pattern="R2_001.fastq", full.names = TRUE))

```

## Assign the filenames for the filtered fastq.gz files and place in filtered subdirectory 
```{r}
sample.names <- sapply(strsplit(basename(fnFs), "_"), `[`, 1)

# We'll still have the "-bac" at the end of the file names. 
#To remove that, we'll use gsub()

sample.names <- gsub("-euc", "", sample.names)

# Let's make sure the previous functions worked

sample.names

#Ok we did it. Now, let's put the files in a subdirectory

filt_path <- file.path(getwd(), "~/sequences/Euc")
filtFs <- file.path("~/sequences/Euc/filtered", paste0(sample.names, "_F_filt.fastq.gz"))
filtRs <- file.path("~/sequences/Euc/filtered", paste0(sample.names, "_R_filt.fastq.gz"))

names(filtFs) <- sample.names
names(filtRs) <- sample.names
``` 


## Let's check the quality of our sequences

```{r}
#First, we can look at some of the Forwards sequences for Filter samples

plotQualityProfile(fnFs[9:18])

#Save as a PDF file

ggsave("sequences/Euc/euc_Figures/Euc_qualityplot_Filtre_Forward.pdf", height = 8, width = 12, dpi = 16, plot = last_plot())

#Then, some of the Forwards sequences for Rock samples 

plotQualityProfile(fnFs[40:50])
ggsave("sequences/Euc/euc_Figures/Euc_qualityplot_Roche_Forward.pdf", height = 8, width = 12, dpi = 16, plot = last_plot())


#Now we can look at some of the Reverse sequences for Filter samples

plotQualityProfile(fnRs[9:18])
ggsave("sequences/Euc/euc_Figures/Euc_qualityplot_Filtre_Reverse.pdf", height = 8, width = 12, dpi = 16, plot = last_plot())


#And finally, some of the Reverse sequences for Rock samples

plotQualityProfile(fnRs[40:50])
ggsave("sequences/Euc/euc_Figures/Euc_qualityplot_Roche_Reverse.pdf", height = 8, width = 12, dpi = 16, plot = last_plot())


```


## Time to trim the reads thanks to the quality graphs (see previous section of code) 
## From the previous section, it looks like the quality diminishes at the 270 mark.
## For the forward sequences, we'll remove the first 17 nucleotides (size of the primer)
## For the reverse, we'll remove the first 21 nucleotides (size of the primer).
## We'll trim the last nucleotides at position 270 (trimming the last 70 nucleotides) to avoid errors
```{r}

out <- filterAndTrim(fnFs, filtFs, fnRs, filtRs, truncLen=c(270,200),
                     maxN=0, maxEE=c(5,5), truncQ=2, trimLeft=c(17,21),
                     compress=TRUE, multithread=TRUE) 
out

```

## Let's now look at the error rate

```{r}

errF <- learnErrors(filtFs, multithread=TRUE, randomize=TRUE)
#101159773 total bases in 399841 reads from 36 samples will be used for learning the error rates.


errR <- learnErrors(filtRs, multithread=TRUE, randomize=TRUE)
#100791141 total bases in 563079 reads from 42 samples will be used for learning the error rates.

#Plot the error rates

plotErrors(errF, nominalQ=TRUE)
dev.print(width = 20, height = 10, pdf,"/home/dpatel/sequences/Euc/euc_Figures/plotErrors.pdf")

```

## Time to dereplicate: removing replicate of identical sequences from each sample

```{r}
derepFs <- derepFastq(filtFs, verbose=TRUE)
derepRs <- derepFastq(filtRs, verbose=TRUE)


# Name the derep-class objects by the sample names

names(derepFs) <- sample.names
names(derepRs) <- sample.names

# Apply the core sample inference algorithm to the filtered and trimmed sequence data
dadaFs <- dada(filtFs, err=errF, pool="pseudo", multithread=TRUE) 
dadaRs <- dada(filtRs, err=errR, pool="pseudo", multithread=TRUE)

# Inspect the returned dada-class object:
dadaFs[[7]]

# Result = dada-class: object describing DADA2 denoising results
# 6 sequence variants were inferred from 419 input unique sequences.

```



## We're going to create the ASVs 
```{r}

# First, we merge the paired reads

mergers <- mergePairs(dadaFs, filtFs, dadaRs, filtRs)

# Before continuing, let's look at the data.frame (first sample only)

head(mergers[[1]])

# Construct an amplicon sequence variant table (ASV) table and check the dimension of it

seqtab <- makeSequenceTable(mergers)
dim(seqtab)

#[1]    62 868

# Inspect distribution of sequence lengths

table(nchar(getSequences(seqtab)))
#The majority of the sequences are 253 nucleotides


# Remove chimeras 

seqtab.nochim <- removeBimeraDenovo(seqtab, method="consensus", multithread=TRUE)


#View dimension of your matrices 

dim(seqtab.nochim)
sum(seqtab.nochim)/sum(seqtab)

#[1]   62 574
#[1] 0.895732
``` 

## Track reads through the pipeline
```{r}
# Obtain the count of how many sequences were deleted at each steps. 

getN <- function(x) sum(getUniques(x))
track <- cbind(out, sapply(dadaFs, getN), sapply(dadaRs, getN), sapply(mergers, getN), rowSums(seqtab.nochim))
colnames(track) <- c("input", "filtered", "denoisedF", "denoisedR", "merged", "nonchim")
rownames(track) <- sample.names
track
``` 

## Time to asing taxonomy to our ASVs

```{r}

# Identifying the taxonomy

taxa <- assignTaxonomy(seqtab.nochim, "~/silva_132.18s.99_rep_set.dada2.fa.gz?download=1", multithread=TRUE, tryRC=TRUE)

#taxa.sp <- addSpecies(taxa,  "~/silva_species_assignment_v138.1.fa.gz?download=1", allowMultiple = TRUE, tryRC = TRUE)

# Removing sequence rownames for display only

taxa.print <- taxa 
rownames(taxa.print) <- NULL
head(taxa.print, n=20)

View(taxa)


```


## Save as .CSV files and save the sequence table

```{r}

# For the taxonomy table

write.csv(as.data.frame(taxa.print), file = file.path("~/sequences/Euc/euc_csv/EUC_S1_taxonomy_table.csv")) 

# For the ASV table

write.csv(as.data.frame(seqtab.nochim),file = file.path("~/sequences/Euc/euc_csv/ASV_euc1_table.csv")) 
write.csv(as.data.frame(t(seqtab.nochim)),file = file.path("~/sequences/Euc/euc_csv/ASV_E1_t_table.csv"))

#Save sequence table and taxonomic annotations to individual files

saveRDS(seqtab.nochim, file = file.path("~/sequences/euc1_seqtab_nochim.rds"))
saveRDS(taxa, file = file.path("~/sequences/euc1_taxa_sp.rds"))

```


## Loading libraries needed for the cleaning and summarizing of our DADA2 results

```{r}
# Let's open all the data we'll need  

library(picante)
library(vegan)
library(ggplot2)
library(reshape2)
library(ggpubr)
library(DESeq2)
library(pheatmap)
library(phyloseq)

```

## Let's also get our DADA2 results

```{r}
# load sequence table (nonchimeric ASV abundances in samples)

seqtab.nochim <- readRDS("/home/dpatel/sequences/Euc/euc1_seqtab_nochim.rds")


# load taxonomic annotations (taxonomic ID of each ASV)

taxa.sp <- readRDS("/home/dpatel/sequences/Euc/euc1_taxa_sp.rds")

```

## Let's create a community and taxonomy object

```{r}
# community object from nonchimeric ASV sequence table

comm <- seqtab.nochim


# taxonomy object - make sure order of ASVs match between community and taxonomy

taxo <- taxa.sp[colnames(comm),]

```



## Doing some cleaning

```{r}
# First, let's make sure we only have bacteria phylums in our taxonomy
table(taxo[,"Kingdom"])


#Let's also put our taxo as a data frame to be able to use phyloseq.
#We can also use this opportunity to put "NAs".

taxid <- as.data.frame(t(taxo))
taxid[] <- lapply(taxid, as.character)
taxid2<- tidyr::fill(taxid, colnames(taxid),.direction = "down")
taxid2<- sapply(taxid2, function(x){paste0("unclassified_", x)})
taxid[is.na(taxid)] <- taxid2[is.na(taxid)]
taxid <- t(taxid)
taxid[ taxid == "unclassified_NA" ] <- NA

# Remove unclassified ASV from the taxonomy matrix then from the ASV matrix

taxid <-subset(as.data.frame(taxid), Kingdom =="Eukaryota")
comm <- seqtab.nochim[,colnames(comm) %in% rownames(taxid)]


```


## Looking at our data.

```{r}
# Number of reads per sample
rowSums(comm)

# visualize log10 number of reads per sample

hist(log10(rowSums(comm)))
#We have a lot of reads in some samples and very few reads in other (most likely the control samples)

# log10 of number of reads per ASV
hist(log10(colSums(comm)))

# Looks like we're going to have a couple of dominant ASVs in our samples

```


## Visualize community composition

```{r}
# PCA on Hellinger-transformed community data
comm.pca <- prcomp(decostand(comm,"hellinger"))

# plot ordination results
ordiplot(comm.pca, display="sites", type="text",cex=1.0)

# Our samples (except the F-R-J1) are all seperate from our Controls (CRTL) and negatives. Some of the CRTL are pretty close to the Tourb


# Surface plot - plot number of sequences/sample (library size)

ordisurf(comm.pca, rowSums(comm), bubble=TRUE, cex=2,
         main="Library size (sequences/sample)")
dev.print(width = 10, height = 8, pdf,"/home/dpatel/sequences/Euc/euc_Figures/seq_per_sample.pdf")

# Most samples are between 12 000 and 16 000 sequences. There are a couples of samples at around 0 sequence. 

```


## Check negative controls

```{r}
# Abundance of ASVs in negative controls

comm['BlancKit',][comm['BlancKit',]>0]
comm['CRTL-B4B',][comm['CRTL-B4B',]>0]
comm['CRTL-B5H',][comm['CRTL-B5H',]>0]
comm['CRTL-J1B',][comm['CRTL-J1B',]>0]
comm['CRTL-J2M',][comm['CRTL-J2M',]>0]
comm['CRTL-R3H',][comm['CRTL-R3H',]>0]
comm['CRTL-R6M',][comm['CRTL-R6M',]>0]


```


Create phyloseq containing the phylogenetic tree and the metadata 
```{r}
## PHYLOSEQ OBJECT 
library(Biostrings); packageVersion("Biostrings")
library(decontam)
library(phyloseq)

# load metadata
metadata <- read.csv("/home/dpatel/metadata_Bioreacteur1.csv", header=TRUE, sep=",", row.names = 1)

# inspect metadata
head(metadata)

#Phyloseq 

ps <- phyloseq(otu_table(t(comm), taxa_are_rows=TRUE), tax_table(as.matrix(taxid)), sample_data(metadata))


# Add sequence table 
dna <- Biostrings::DNAStringSet(taxa_names(ps))
names(dna) <- taxa_names(ps)
ps <- merge_phyloseq(ps, dna)

# Make ASV ID shorter 
taxa_names(ps) <- paste0("ASV", seq(ntaxa(ps)))

```

## Remove the negative controls (decontam)

```{r}

sample_data(ps)$is.neg <- sample_data(ps)$SampleType == "Controle"
contamdf.prev <- isContaminant(ps, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev$contaminant)
ps.noncontam <- prune_taxa(!contamdf.prev$contaminant, ps)
final.ps=subset_samples(ps.noncontam, !SampleType=="Controle")

#Save the ASVs, taxa and metadata

write.csv(as.data.frame(as(tax_table(final.ps), "matrix")), file = "/home/dpatel/sequences/Euc/euc_csv/noncontam_taxa.csv") 
write.csv(as.data.frame(as(otu_table(final.ps), "matrix")),file = "/home/dpatel/sequences/Euc/euc_csv/noncontam_asv.csv")
write.csv(as.data.frame(as(sample_data(final.ps), "matrix")), file="/home/dpatel/sequences/Euc/euc_csv/noncontam_meta.csv")

```



## Visualize the new community composition

```{r}

# Load our "noncontam" files

noncontam_taxa <- read.csv("/home/dpatel/sequences/Euc/euc_csv/noncontam_taxa.csv", header = TRUE, sep =",", row.names = 1)
noncontam_asv <- read.csv("/home/dpatel/sequences/Euc/euc_csv/noncontam_asv.csv", header = TRUE, sep =",", row.names = 1)
noncontam_meta <- read.csv("/home/dpatel/sequences/Euc/euc_csv/noncontam_meta.csv", header=TRUE, sep=",", row.names = 1)

# Puttingg our ASVs and taxonomy into another object 

noncontam_comm <- t(as.data.frame(noncontam_asv))
noncontam_taxo <- as.data.frame(noncontam_taxa)[colnames(noncontam_comm),]

# PCA on Hellinger-transformed community data
noncontam.pca <- prcomp(decostand(noncontam_comm,"hellinger"))
ordiplot(noncontam.pca, display="sites", type="text",cex=1.0)
dev.print(width = 10, height = 8, pdf,"/home/dpatel/sequences/Euc/euc_Figures/PCA.pdf")
# Our samples are all pretty close from each other


```



## Making rarefaction curves 

```{r}

# What is the smallest number of sequences/sample for subset of samples?

min(rowSums(noncontam_comm))
rowSums(noncontam_comm)

# Rarefaction curve for subset of samples + saving them

rarecurve(noncontam_comm, step=200, label=TRUE)
dev.print(width = 24, height = 10, pdf,"/home/dpatel/sequences/Euc/euc_Figures/rarefaction_curve_all.pdf")

rarecurve(noncontam_comm, step=200, label=TRUE, xlim=c(0,5000))
dev.print(width = 24, height = 10, pdf,"/home/dpatel/sequences/Euc/euc_Figures/rarefaction_curve_short.pdf")


minTotRelAbun = 5e-5
x = taxa_sums(final.ps)
keepTaxa =  (x / sum(x)) > minTotRelAbun
prunedSet = prune_taxa(keepTaxa, final.ps)
sample_sums(prunedSet)
sample_sums(final.ps)

rarecurve(t(otu_table(final.ps)), step=50, cex=0.5)
ps_raref <-rarefy_even_depth(prunedSet, 3000, rngseed=112) 

#Save asv, taxa, metadata rarefied
write.csv(as.data.frame(as(tax_table(ps_raref), "matrix")), file = "/home/dpatel/sequences/Euc/euc_csv/rarefied_taxa.csv") 
write.csv(as.data.frame(as(otu_table(ps_raref), "matrix")),file = "/home/dpatel/sequences/Euc/euc_csv/rarefied_asv.csv")
write.csv(as.data.frame(as(sample_data(ps_raref), "matrix")), file="/home/dpatel/sequences/Euc/euc_csv/rarefied_meta.csv") 

```


## Preparing for the plots (BONNE)

```{r}
#install.packages("randomcoloR")
library(plyr)
library(dplyr)
library(phyloseq)
library(randomcoloR)


asv=read.table("/home/dpatel/sequences/Euc/euc_csv/rarefied_asv.csv", sep=",", row.names=1, header=TRUE, check.names=FALSE) 
taxa=read.table("/home/dpatel/sequences/Euc/euc_csv/rarefied_taxa.csv", sep=",", row.names=1, header=TRUE) 
meta=read.table("/home/dpatel/sequences/Euc/euc_csv/rarefied_meta.csv", sep=",", row.names=1, header=TRUE) 

# Merge into phyloseq object 
ps=phyloseq(otu_table(asv, taxa_are_rows=TRUE), tax_table(as.matrix(taxa)), sample_data(meta))

# Select taxonomic rank
taxa_rank <- readline(prompt="Which taxnomic rank do you want to plot, Genus (G) or Class (C) ? ")
if (taxa_rank == "G"){
  taxa_rank="Genus"
} else if (taxa_rank == "C"){
  taxa_rank="Class"
}else {
  print("wake up brain")
}

# Transform counts to relative abundance and agglomerate to the desired taxonomic rank. Melt phyloseq object to dataframe. 

ps_rel_abund=transform_sample_counts(ps, function(x) x/sum(x)) # get abundance in %
glom=tax_glom(ps_rel_abund, taxrank = taxa_rank) # agglomerate taxa at the desired taxonomic level
df=psmelt(glom) # create dataframe from phyloseq object

medians=ddply(df, taxa_rank, function(x) c(median=median(x$Abundance)))
remainder <- medians[medians$median <= 0.001,][[1]] # Find taxons whose rel. abund. is less than 0.1%
df[df[[taxa_rank]] %in% remainder,][[taxa_rank]]="Other (abundance < 0.1%)" # Change their name to "Other (abondance < 0.1%)"

# Create empty list 
list_of_all_taxonomic_rank= list()

```



### Making the plots
```{r}

# Plot with Class: You have to type "C" at the "Select taxonomic rank" part

# We'll start with the Class for our Filtre samples

Class_Filters <- df[(df$SampleType == "Filtre"), ]
Class_Filters$Sample <- gsub("F-", "", Class_Filters$Sample)

Class_Filters$Sample <- factor(Class_Filters$Sample, c("P1","P2","P3","P4","Tourb",
                              "B-J1","B-J3","B-J5","B-J7","B-J9",
                              "J-J1","J-J2","J-J3","J-J5","J-J6","J-J7","J-J8","J-J9","J-J11",
                              "R-J2","R-J3","R-J5","R-J6","R-J7","R-J8","R-J9","R-J10")) 

BAC_Class=ggplot(Class_Filters, aes(x=Sample, weight=Abundance, fill=.data[[taxa_rank]])) +
  geom_bar() +
  labs(y ='Relative abundance (%)', paste(sample), x = "Filter") +
  scale_y_continuous(expand = c(0,0)) +
  theme(text = element_text(size = 20),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale + 
  facet_grid(scales="free", margins = FALSE, drop = F) +
  ggtitle(" ")
  
BAC_Class
ggsave("sequences/Euc/euc_Figures/Euc_Filtre_Class.pdf", width = 24, height = 12, plot = last_plot())


# Plot with Class (Rocks)

Class_Rocks <- df[(df$SampleType == "Roche"),] 
Class_Rocks$Sample <- gsub("R-", "", Class_Rocks$Sample)

Class_Rocks$Sample <- factor(Class_Rocks$Sample, c("B1B","B5B","B7B","B8B","B3M","B4M","B1H","B6H","B7H",
                                                   "J1B","J3B","J7B","J8B","J2H","R5B","R7B","R2M","R3M","R7M","R1H"))

BAC_Class_R=ggplot(Class_Rocks, aes(x=Sample, weight=Abundance, fill=.data[[taxa_rank]])) +
  geom_bar() +
  labs(y ='Relative abundance (%)', paste(sample), x = "Rock") +
  scale_y_continuous(expand = c(0,0)) +
  theme(text = element_text(size = 20),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale + 
  facet_grid(cols = vars(Class_Rocks$Bioreacteur), scales="free", margins = FALSE, drop = F) +
  ggtitle(" ")
  
BAC_Class_R
ggsave("sequences/Euc/euc_Figures/Euc_Roche_Class.pdf", width = 24, height = 12, plot = last_plot())



```



### Making the Genus plots
```{r}

# Merge into phyloseq object 
ps=phyloseq(otu_table(asv, taxa_are_rows=TRUE), tax_table(as.matrix(taxa)), sample_data(meta))

# Select taxonomic rank
taxa_rank <- readline(prompt="Which taxnomic rank do you want to plot, Genus (G) or Class (C) ? ")
if (taxa_rank == "G"){
  taxa_rank="Genus"
} else if (taxa_rank == "C"){
  taxa_rank="Class"
}else {
  print("wake up brain")
}

# Transform counts to relative abundance and agglomerate to the desired taxonomic rank. Melt phyloseq object to dataframe. 

ps_rel_abund=transform_sample_counts(ps, function(x) x/sum(x)) # get abundance in %
glom=tax_glom(ps_rel_abund, taxrank = taxa_rank) # agglomerate taxa at the desired taxonomic level
df=psmelt(glom) # create dataframe from phyloseq object

medians=ddply(df, taxa_rank, function(x) c(median=median(x$Abundance)))
remainder <- medians[medians$median <= 0.001,][[1]] # Find taxons whose rel. abund. is less than 0.1%
df[df[[taxa_rank]] %in% remainder,][[taxa_rank]]="Other (abundance < 0.1%)" # Change their name to "Other (abondance < 0.1%)"


# Create empty list 
list_of_all_taxonomic_rank= list()

# Loop over every sample_type to combine taxons with a relative abundance < 0.1% 
data <- data.frame()
 

for (sample in unique(df$Sample)){
    table=subset(df, Sample==sample)
    medians=ddply(table, taxa_rank, function(x) c(median=median(x$Abundance)))
    remainder <- medians[medians$median <= 0.001,][[1]] # Find taxons whose rel. abund. is less than 0.1%
    table[table[[taxa_rank]] %in% remainder,][[taxa_rank]]="Other (abundance < 0.1%)" # Change their name to "Other (abondance < 0.1%)"
    list_of_all_taxonomic_rank <- append(list_of_all_taxonomic_rank,table[[taxa_rank]])
data = rbind(df,table)

    
taxon_list=as.data.frame(list_of_all_taxonomic_rank)
df_taxon_list=as.data.frame(t(taxon_list))
unique_taxonomic_ranks=as.data.frame(unique(df_taxon_list$V1))
names(unique_taxonomic_ranks)[names(unique_taxonomic_ranks)=="unique(df_taxon_list$V1)"]=as.character(taxa_rank) # Rename column 

}


# Plot with Genus: You have to type "G" at the "Select taxonomic rank" part

# We'll start with our Filtre samples

Genus_Filters <- df[(df$SampleType == "Filtre"), ]
Genus_Filters$Sample <- gsub("F-", "", Genus_Filters$Sample)

Genus_Filters$Sample <- factor(Genus_Filters$Sample, c("P1","P2","P3","P4","Tourb",
                              "B-J1","B-J3","B-J5","B-J7","B-J9",
                              "J-J1","J-J2","J-J3","J-J5","J-J6","J-J7","J-J8","J-J9","J-J11",
                              "R-J2","R-J3","R-J5","R-J6","R-J7","R-J8","R-J9","R-J10")) 

EUC_Genus=ggplot(Genus_Filters, aes(x=Sample, weight=Abundance, fill=.data[[taxa_rank]])) +
  geom_bar() +
  labs(y ='Relative abundance (%)', paste(sample), x = "Filter") +
  scale_y_continuous(expand = c(0,0)) +
  theme(text = element_text(size = 20),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale + facet_grid( scales="free", margins = FALSE, drop = FALSE) +
  ggtitle(" ")
  
EUC_Genus
ggsave("sequences/Euc/euc_Figures/Euc_Filtre_Genus.pdf", width = 24, height = 12, plot = last_plot())


# Plot with Genus (Rocks)

Genus_Rocks <- df[(df$SampleType == "Roche"),] 
Genus_Rocks$Sample <- gsub("R-", "", Genus_Rocks$Sample)

Genus_Rocks$Sample <- factor(Genus_Rocks$Sample, c("B1B","B5B","B7B","B8B","B3M","B4M","B1H","B6H","B7H",
                                                   "J1B","J3B","J7B","J8B","J2H","R5B","R7B","R2M","R3M","R7M","R1H"))

EUC_Genus_R=ggplot(Genus_Rocks, aes(x=Sample, weight=Abundance, fill=.data[[taxa_rank]])) +
  geom_bar() +
  labs(y ='Relative abundance (%)', paste(sample), x = "Rock") +
  scale_y_continuous(expand = c(0,0)) +
  theme(text = element_text(size = 20),axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  my_scale + 
  facet_grid(cols = vars(Genus_Rocks$Bioreacteur), scales="free", margins = FALSE, drop = F) +
  ggtitle(" ")
  
EUC_Genus_R
ggsave("sequences/Euc/euc_Figures/Euc_Roche_Genus.pdf", width = 24, height = 12, plot = last_plot())



```


## SET COLORS
```{r}
# Assign colors to every unique taxonomic_ranks 

#my_scale <- scale_fill_manual(name=as.character(taxonomic_rank), values = palette, na.translate=FALSE, drop=TRUE, limits = force) # Recreate color fill 


# To create a new color range
n <- nrow(unique_taxonomic_ranks) + 20
palette <- distinctColorPalette(n)
unique_taxonomic_ranks[[taxa_rank]]=factor(unique_taxonomic_ranks[[taxa_rank]]) 
names(palette) <- levels(unique_taxonomic_ranks[[taxa_rank]]) 
my_scale <- scale_fill_manual(name=as.character(taxa_rank), values = palette, na.translate=FALSE, drop=TRUE, limits = force)

# Saving the color palette for future use 
#saveRDS(palette, "/home/dpatel/sequences/Arc/colors_class.rds")

```





# Let's just see the effect of the rarefaction.

```{r}

set.seed(0)

# Randomly rarefy samples

comm_rarfy <- rrarefy(noncontam_comm, sample = min(rowSums(noncontam_comm)))

# Remove any ASVs whose abundance is 0 after rarefaction

comm_rarfy <- comm_rarfy[,colSums(comm_rarfy)>1]

# Match ASV taxonomy to rarefied community

taxo_rarfy <- taxid[colnames(comm_rarfy),]


# Rarefaction is not expected to have a major influence on the diversity (as rarefaction curves have reached the plateau).

richness_raw <- rowSums((noncontam_comm>0)*1)

richness_rarfy <- rowSums((comm_rarfy>0)*1)

plot(richness_rarfy~richness_raw,xlab='number of ASVs in raw data',ylab='number of ASVs in rarefied data')
reg <- lm(richness_rarfy~richness_raw)
abline(reg, col = "red")


```


