---
title: "Diversite_beta_arc"
author: "Divya"
date: "2023-05-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list=ls())
```


### Load libraries
```{r message=TRUE}

library(vegan)
library(tidyr)
library(dplyr)
library(ggplot2)

```


### Loading all files needed here
```{r}

# Load ASV table and metadatas
ASV_table <- t(read.table("~/sequences/Arc/arc_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Arc/arc_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)

```


### Doing some cleaning in the dataframes
```{r}
# Remove the mysterious "is.neg" column and remove the Rugosite1 and Rugosite2 in Metadata since we don't need them
metadata <- metadata %>% select(-is.neg,-Rugosite1,-Rugosite2)


# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"

```


### Correlation between our environemental factors and our beta diversity
```{r}
# Let's start by doing a dissimilarity matrix (Bray-Kurtis)
beta_dist <- vegdist(ASV_table, method = "bray")
write.csv(as.matrix(beta_dist), "/home/dpatel/sequences/Arc/arc_csv/dist_matrix.csv")
beta_dist <- read.table("/home/dpatel/sequences/Arc/arc_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = NULL)
colnames(beta_dist) <- c("Nom", beta_dist$X)

# Now we have to join the metadata dataframe with the matrix
meta_distance <- inner_join(metadata, beta_dist, by = "Nom")

all_dist <- meta_distance %>% select(all_of(meta_distance$Nom)) %>% as.dist()

# Run the adonis2 function and look at the results
Corr <- adonis2(all_dist ~ Bioreacteur, data = meta_distance, by="margin")
print(Corr)

```



### Plot a PCoA
```{r}
# Load the dissimilarity matrix
beta_dist <- read.table("/home/dpatel/sequences/Arc/arc_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = 1)

# Perform PCoA analysis using the "cmdscale" function
PCOA <- cmdscale(beta_dist, eig=TRUE, add=TRUE)

# Extract the coordinate results from the PCOA
pcoa_df <- data.frame(PCoA1 = PCOA$points[,1], PCoA2 = PCOA$points[,2], Sample = row.names(beta_dist))

pcoa_df$SampleType <- ifelse(grepl("^F", pcoa_df$Sample), "Filter",
                        ifelse(grepl("^R", pcoa_df$Sample), "Rock",
                               ifelse(grepl("^T|^P", pcoa_df$Sample), "Field", "Other")))

# Remove "R." and "F." from the sample names in the PCoA data frame and change the J of Jour to D (day)
pcoa_df$Sample <- gsub("^R\\.|^F\\.", "", pcoa_df$Sample)
pcoa_df$Sample <- gsub(".J", "-D", pcoa_df$Sample)

# Extract numbers from the end of Sample names and multiply by 2
pcoa_df <- pcoa_df %>%
  mutate(Sample = ifelse(Sample %in% c("P1", "P2", "P3", "P4"), Sample, str_replace(Sample, "(\\d+)$", function(match) as.character(2 * as.numeric(match)))))


# Create a new column in the PCoA data frame that indicates the shape of each sample
pcoa_df$Shape <- ifelse(grepl("^B", pcoa_df$Sample), "circle",
                        ifelse(grepl("^J", pcoa_df$Sample), "diamond",
                         ifelse(grepl("^R", pcoa_df$Sample), "square", "triangle")))

pcoa_df$Sample<- gsub("Tourb", "Peat", pcoa_df$Sample)
percent_explained=100*PCOA$eig/sum(PCOA$eig)
reduced_percent=format(round(percent_explained[1:2], digits=2), nsmall=1, trim=TRUE) # reduce number of digits

# Visualize the PCoA plot using ggplot
p <- ggplot(pcoa_df, aes(x = PCoA1, y = PCoA2, color = SampleType, shape = Shape, label = Sample)) +
  geom_point(size = 3.5) +
  xlab(paste0("PCoA 1 (", reduced_percent[1], "% of variation)")) +
  ylab(paste0("PCoA 2 (", reduced_percent[2], "% of variation)")) +
  geom_text(hjust = 0.5, vjust = -0.5) +
  theme(legend.position = "right") +  # move legend to the right
  scale_color_manual(values = c("#116530","#21B6A8","#A45C40"), # Set colors
                     name = "Sample Type") +  # define legend title and labels
  scale_shape_manual(values = c(16, 17, 15, 18),  # Set the shape scheme
                      labels = c("Bioreactor 1", "Bioreactor 2", "Bioreactor 3", "Field"), # Set the labels
                     name = "Bioreactor")

# Show the plot
p + theme(text = element_text(size = 20), legend.text = element_text(size = 16))  

ggsave("/home/dpatel/sequences/Arc/arc_Figures/PCOA_beta1_arc.tiff", width = 8, height = 6, plot = last_plot())

```


### PERMANOVA FOR FILTERS ###

# After taking a look at the PCoA, we'll separate the data based on the Jours. 
# So the days will be J1-J6 and J7-J9
```{r}
rm(list=ls())

# Load ASV table and metadatas
ASV_table <- t(read.table("~/sequences/Arc/arc_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Arc/arc_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)
beta_dist <- read.table("/home/dpatel/sequences/Arc/arc_csv/dist_matrix.csv", header=TRUE, sep=",", row.names = NULL)
colnames(beta_dist) <- c("Nom", beta_dist$X)
```

### Doing some cleaning in the dataframes
```{r}
# Remove the mysterious "is.neg" column and remove the unecessary column in metadata 
metadata <- metadata %>% select(-is.neg, -Rugosite1, -Rugosite2, -concADN)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"
```

### Doing the PERMANOVA (SEPARATE BASED ON JOURs)
```{r}
# Separate the Filtres based on the Jours 
set.seed(123)
meta16  <- metadata[metadata$Jour %in% c("2", "4", "6","10", "12"), ]
meta710 <- metadata[metadata$Jour %in% c("14", "16", "18", "20"), ]

# Now we have to join the metadata dataframes with the matrix and keep the lower half of the dist
meta16_dist <- inner_join(meta16, beta_dist, by = "Nom")
dist16 <- meta16_dist %>% select(all_of(meta16_dist$Nom)) %>% as.dist()

meta710_dist <- inner_join(meta710, beta_dist, by = "Nom")
dist710 <- meta710_dist %>% select(all_of(meta710_dist$Nom)) %>% as.dist()


# Run the adonis2 function and look at the results
Corr16 <- adonis2(dist16 ~ Bioreacteur + Jour, data = meta16_dist, by="margin")
print(Corr16)


Corr710 <- adonis2(dist710 ~ Bioreacteur + Jour, data = meta710_dist, by="margin")
print(Corr710)


```

### Doing the PERMANOVA (ALL FILTERS COMBINED)
```{r}
# Keeping the Filtres only
meta_filter <- metadata[(metadata$SampleType == "Filtre"), ]
meta_filter <- meta_filter[!(meta_filter$Bioreacteur == "Terrain"), ]

# Now we have to join the metadata dataframes with the matrix and keep the lower half of the dist
meta_dist <- inner_join(meta_filter, beta_dist, by = "Nom")
dist <- meta_dist %>% select(all_of(meta_dist$Nom)) %>% as.dist()

# Run the adonis2 function and look at the results
Corr <- adonis2(dist ~ Bioreacteur + Jour, data = meta_dist, by="margin")
print(Corr)
#             Df  SumOfSqs    R2       F     Pr(>F)    
# Bioreacteur  2  0.16114   0.13059  1.5804  0.041 *  
# Jour         1  0.32232   0.26121  6.3224  0.001 ***
# Residual    15  0.76470   0.61972                  
# Total       18  1.23395   1.00000       

```






###### Doing a dbRDA ########

# For the dbRDA, The ASV dataframe should have samples as rows and ASVs as columns
# As for the metadata dataframe, it should have samples as rows and metadata variables as columns.


### FILTERS ONLY

### Loading dataframes and doing some cleaning
```{r}
rm(list=ls())

# Load the ASV table and the metadata
ASV_table <- t(read.table("~/sequences/Arc/arc_csv/rarefied_asv.csv", header = TRUE, row.names = 1, sep = ","))
metadata <- read.table("~/sequences/Arc/arc_csv/rarefied_meta.csv", header=TRUE, sep=",", row.names = NULL)

# Remove the mysterious "is.neg" column and remove the Rugosite1 and Rugosite2 because we don't need them
metadata <- metadata %>% select(-is.neg,-Rugosite1,-Rugosite2)

# Since our sample names differ in our ASV and Metadata table, we'll have to replace the "-" to "."
# We'll also change the name of the first column to "Nom"
metadata$X <- gsub("-", ".", metadata$X)
colnames(metadata)[1] <- "Nom"


# Taking all my Filter data from metadata
Meta_filters_all <- metadata[(metadata$SampleType == "Filtre"), ]

# Merging the geochemical data to the reste of our metadata.
Analyse <- read.table("/home/dpatel/Analyse_geo_merged.csv", header=TRUE, sep=",", row.names = NULL)
Meta_filters_all <- merge(Meta_filters_all, Analyse, by = c("Bioreacteur", "Jour"), sort = TRUE)

# Change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
Meta_filters_all$Bioreacteur <- ifelse(grepl("^B", Meta_filters_all$Bioreacteur), "Bio1",
                        ifelse(grepl("^J", Meta_filters_all$Bioreacteur), "Bio2",
                         ifelse(grepl("^R", Meta_filters_all$Bioreacteur), "Bio3", "Field")))


# Taking the ASV table and keeping the same samples as in the Filtres_noT_geo metadata.
# We'll also transform the raw ASV data into relative abundance 
ASV_table <- ASV_table[Meta_filters_all$Nom, ]
ASV_table <- decostand(ASV_table, method = "hellinger")

# Removing "F." from the sample names for the ASV table and the Meta_filters_all dataframe 
rownames(ASV_table) <- gsub("F\\.", "", rownames(ASV_table))
Meta_filters_all$Nom <- gsub("F\\.", "", Meta_filters_all$Nom)

# Change the J of Jour to D (day)
Meta_filters_all$Nom <- gsub(".J", "-D", Meta_filters_all$Nom)


# Extract numbers from the end of Sample names and multiply by 2
Meta_filters_all <- Meta_filters_all %>%
mutate(Nom = str_replace(Nom, "(\\d+)$", function(match) as.character(2 * as.numeric(match))))

```

### Doing the Distance matrix and dbRDA 
```{r}

beta_dist <- vegdist(ASV_table, method = "bray")

dbRDA_result <- capscale(beta_dist ~ Moy_DIC + Moy_DOC + Moy_NO2 + Moy_NOx + Moy_NH4 + Moy_NO3 , data = Meta_filters_all)

```

### Looking at the plot
```{r}
# Doing a simple plot
plot(dbRDA_result)
```

# Doing some statistical test
```{r}
anova(dbRDA_result) 

# test axes for significance
anova(dbRDA_result, by="axis", perm.max=1000) 

# test for significant variables from the matadata
anova(dbRDA_result, by="terms", permu=1000) 

summary(dbRDA_result)

```

### Ok, Let's see if we can simplify our dbRDA
```{r}
#So step() is going to find the simplest models we can have: taking away variables that aren't too important
mod1 <- step(dbRDA_result, scope = formula(dbRDA_result), criteria = "AIC", test = "perm", direction = "both") 
summary(mod1)

mod2 <- capscale(beta_dist ~ Moy_DIC + Moy_DOC, data = Meta_filters_all)

set.seed(123)
# Looking for some collinearity
vif.cca(mod2)

#Doing some statistical analysis
anova(mod2) 

# test axes for significance
anova(mod2, by="axis", perm.max=1000) 

# test for significant variables from the metadata
anova(mod2, by="terms", permu=1000) 

#summary(mod2)

plot(mod2)
```

### Plotting
```{r}

# Extract the site scores from dbRDA result
site_scores <- scores(mod2, display = "sites")
site_scores_df <- as.data.frame(site_scores)

# Extract the variable scores from dbRDA result
variable_scores <- scores(mod2, display = "bp")
variable_scores_df <- as.data.frame(variable_scores)

# Get the dimension names from the site scores
dim_names <- colnames(site_scores_df)

# Calculate the eigenvalues of the dbRDA result
eigenvals <- eigenvals(mod2)

# Calculate the percentage of variance explained by each dimension
var_explained <- eigenvals / sum(eigenvals) * 100

# Create a dataframe combining site scores and metadata
plot_data <- cbind.data.frame(site_scores_df, Meta_filters_all)

dbRDAfilter <-ggplot() +
  geom_point(data = plot_data, aes(x = CAP1, y = CAP2, color = Bioreacteur), size = 2) + #adding the points of our sample
  scale_color_manual(values = c("#21B6A8","#A45C40","#116530"), name = "Bioreactor",) + 
  geom_segment(data = variable_scores_df, aes(x = 0, y = 0, xend = CAP1, yend = CAP2), 
               arrow = arrow(length = unit(0.15, "cm")), color = "black") + #adding the arrows for the environmental variables
  xlab(paste(dim_names[1], " (", round(var_explained[1], 2), "% )")) + #naming the x axis
  ylab(paste(dim_names[2], " (", round(var_explained[2], 2), "% )")) + #naming the y axis
  geom_hline(yintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_vline(xintercept = 0, linetype = "solid", color = "gray") + #adding dashed lines on the x axis
  geom_text(data = plot_data, aes(x = CAP1, y = CAP2, label = Nom), hjust = 0.5, vjust = -0.5, size = 3.5) + #adding the sample name
  theme_minimal() +
  theme(text = element_text(size = 15)) 

dbRDAfilter
ggsave("/home/dpatel/sequences/Arc/arc_Figures/Arc_dbRDA_filter.tiff", width = 6, height = 4, plot = last_plot())

```
