---
title: "Test Venn Diagram"
author: "Divya"
date: "2023-04-18"
output: html_document
---
#### ---------- BACTERIA ---------- ####
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries 
library(ggplot2)
library(dplyr)
library(phyloseq)
library(ggvenn)
library(tibble)
library(VennDiagram)
library(plyr)
library(gplots)
library(ggVennDiagram)
```


### Loading all files needed here 
```{r}
rm(list=ls())
asv <- read.table("~/sequences/Bac/bac_csv/rarefied_asv.csv", sep=",", row.names=1, header=TRUE, check.names=FALSE) 
asv <- data.frame(t(asv))
meta <- read.table("~/sequences/Bac/bac_csv/rarefied_meta.csv", sep=",", row.names=1, header=TRUE) 
meta <- meta[1:3] # Keep only first 3 columns 
df <- merge(meta,asv, by=0)

```

### Working on the dataframes
```{r}
# Remove the Terrain data
df <- df[!(df$Bioreacteur == "Terrain"), ]

# Set the sample names as the Row.names
row.names(df) <- df$Row.names 

# Removing unecessary columns
df <- df %>% select(-Row.names,-Bioreacteur,-Jour) 

# Aggregate based on the SampleType
agg <- aggregate(x = df[, colnames(df) !="SampleType"], 
                by = list(df$SampleType),
                FUN = sum)

# Set first column as row names and remove first column
row.names(agg) <- agg$Group.1 
agg <- agg[-c(1)]

# Translation of the agg dataframe
agg <- as.data.frame(t(agg))

```

### Putting the Genus in the dataframe
```{r}
taxa = read.table("/home/dpatel/sequences/Bac/bac_csv/rarefied_taxa.csv", sep=",", row.names=1, header=TRUE) 
taxa = taxa[6] # Keep only the Genus column

# Merge taxa so we can have the Genus with the proper ASV.
genus <- merge(agg, taxa, by=0)

# Replace values greater than 0 with values from "Genus" column
genus$Filtre[genus$Filtre > 0] <- genus$Genus[genus$Filtre > 0]
genus$Filtre[genus$Filtre == 0] <- NA
genus$Roche[genus$Roche > 0] <- genus$Genus[genus$Roche > 0]
genus$Roche[genus$Roche == 0] <- NA

# Remove "Row.name" and "Genus" columns
genus <- genus[, !(colnames(genus) %in% c("Row.names", "Genus"))]

# Remove rows with NA values in both "Filtre" and "Roche" columns
genus <- genus[!(is.na(genus$Filtre) & is.na(genus$Roche)), ]

# Remove duplicates in "Filtre" and "Roche" columns
genus <- genus[!duplicated(genus[c("Filtre", "Roche")]), ]

```

### Plotting the Venn diagram (WITHOUT the Terrain data)
```{r}
library(VennDiagram)
venn <- Venn(genus)
data <- process_data(venn)

# Create new column with percentage 
data@region=data@region %>% add_column(percent=data@region$count/sum(data@region$count)*100) 
# Round to no digits
data@region$percent = round(data@region$percent, digits=1) 
# Add percentage sign (%)
data@region$percent=paste0(data@region$percent,"%")  

# Add a new column to your data with the fill colors
data@region <- data@region %>% add_column(fill = c("#BDF0E2", "#D8CFEE", "#C7CBE6"))


# Plot the Venn diagram using ggplot2
ggplot() +
  geom_sf(data = venn_region(data), aes(fill = fill), show.legend = FALSE) + 
  scale_fill_identity() +
  geom_sf(data = venn_setedge(data), size = 10) +   
  geom_sf_text(aes(label = c("Planktonic ","Sessile")), data = venn_setlabel(data), size = 5) +
  geom_sf_label(aes(label = percent), data = venn_region(data), alpha = 0, label.size = NA, size = 5) +
  theme_void()

ggsave("/home/dpatel/sequences/Bac/bac_Figures/Veen_Genus_Bac.jpeg", width = 4, height = 4, plot = last_plot())


```

### Checking which genus is in which category
```{r}
list_genus <- as.list(genus)
tmp <- venn(list_genus)
attr(tmp, "intersections")
```




#### ---------- EUCARYOTES ---------- ####

### Loading all files needed here 
```{r}
rm(list=ls())
asv=read.table("~/sequences/Euc/euc_csv/rarefied_asv.csv", sep=",", row.names=1, header=TRUE, check.names=FALSE) 
tasv=data.frame(t(asv))
meta=read.table("~/sequences/Euc/euc_csv/rarefied_meta.csv", sep=",", row.names=1, header=TRUE) 
meta=meta[1:3] # Keep only first 3 columns 
df=merge(meta,tasv, by=0)

```


### Working on the dataframes
```{r}
# Remove the Terrain data
df <- df[!(df$Bioreacteur == "Terrain"), ]

# Set the sample names as the Row.names
row.names(df) <- df$Row.names 

# Removing unecessary columns
df <- df %>% select(-Row.names,-Bioreacteur,-Jour) 

# Aggregate based on the SampleType
agg <- aggregate(x = df[, colnames(df) !="SampleType"], 
                by = list(df$SampleType),
                FUN = sum)

# Set first column as row names and remove first column
row.names(agg) <- agg$Group.1 
agg <- agg[-c(1)]

# Translation of the agg dataframe
agg <- as.data.frame(t(agg))
```

### Putting the Genus in the dataframe
```{r}
taxa = read.table("/home/dpatel/sequences/Euc/euc_csv/rarefied_taxa.csv", sep=",", row.names=1, header=TRUE) 
taxa = taxa[6] # Keep only the Genus column

# Merge taxa so we can have the Genus with the proper ASV.
genus <- merge(agg, taxa, by=0)

# Replace values greater than 0 with values from "Genus" column
genus$Filtre[genus$Filtre > 0] <- genus$Genus[genus$Filtre > 0]
genus$Filtre[genus$Filtre == 0] <- NA
genus$Roche[genus$Roche > 0] <- genus$Genus[genus$Roche > 0]
genus$Roche[genus$Roche == 0] <- NA

# Remove "Row.name" and "Genus" columns
genus <- genus[, !(colnames(genus) %in% c("Row.names", "Genus"))]

# Remove rows with NA values in both "Filtre" and "Roche" columns
genus <- genus[!(is.na(genus$Filtre) & is.na(genus$Roche)), ]

# Remove duplicates in "Filtre" and "Roche" columns
genus <- genus[!duplicated(genus[c("Filtre", "Roche")]), ]

```

### Plotting the Venn diagram (WITHOUT the Terrain data)
```{r}
venn <- Venn(genus)
data <- process_data(venn)

# Create new column with percentage 
data@region=data@region %>% add_column(percent=data@region$count/sum(data@region$count)*100) 
# Round to no digits
data@region$percent = round(data@region$percent, digits=1) 
# Add percentage sign (%)
data@region$percent=paste0(data@region$percent,"%")  

# Add a new column to your data with the fill colors
data@region <- data@region %>% add_column(fill = c("#BDF0E2", "#D8CFEE", "#C7CBE6"))


# Plot the Venn diagram using ggplot2
ggplot() +
  geom_sf(data = venn_region(data), aes(fill = fill), show.legend = FALSE) + 
  scale_fill_identity() +
  geom_sf(data = venn_setedge(data), size = 10) +   
  geom_sf_text(aes(label = c("Planktonic ","Sessile")), data = venn_setlabel(data), size = 5) +
  geom_sf_label(aes(label = percent), data = venn_region(data), alpha = 0, label.size = NA, size = 5) +
  theme_void()

#ggsave("/home/dpatel/sequences/Euc/euc_Figures/Veen_Genus_Euc.jpeg", width = 4, height = 4, plot = last_plot())
#ggsave("/home/dpatel/sequences/Euc/euc_Figures/Veen_Genus_Euc.jpeg", plot=plot, width=16.5, height=16.5,units="cm")

list_genus <- as.list(genus)
tmp <- venn(list_genus)
attr(tmp, "intersections")

```

### Checking which genus is in which category
```{r}
list_genus <- as.list(genus)
tmp <- venn(list_genus)
attr(tmp, "intersections")
```

