---
title: "STAT"
author: "Divya"
date: "2023-01-24"
output:
  html_document: default
  word_document: default
---


#### Open the libraries we'll need
```{r setup}
knitr::opts_chunk$set(echo = FALSE)
rm(list=ls())
library(MASS)
library(tidyr)
library(stats)
library(ggplot2)
library(dplyr)
```


### Open the .csv file 
```{r echo=TRUE}

Analyse <- read.csv("Analyse_geo_Test1.csv", header = TRUE, sep = ",", dec = "." )

head(Analyse)
```




#_________________________________________|DIC|_________________________________________#

### Create a dataframe with the mean and sd of DIC based on Bioreactor and the day (Jour) 
```{r echo=TRUE}
df <- data.frame()

for(i in unique(Analyse$Jour)) {
  
  subset(Analyse, Jour  == i) -> data_Jour
  
  aggregate(DIC.mg.L. ~ Bioreacteur, data = data_Jour, mean, na.rm = TRUE) -> DIC_moy
  
  aggregate(DIC.mg.L. ~ Bioreacteur, data = data_Jour, sd, na.rm = TRUE) -> DIC_sd
  
  df <- rbind(df,data.frame(DIC_moy,DIC_sd))
  
}
print(df)


#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order + Remove the Tourb and Puit row.
#### We should also have to make sure that the data got rearranged and that the Tourb and Puit row got deleted.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))
df <- df[-c(31,32),]
tail(df, 20)


#### Let's rename the Bioreacteur.1 column as "Jour" since that column is pretty useless for us.
#### While we're at it, we can also change the name of our other column so that they represent the values properly.
#### Plus, let's also make sure our changes worked
colnames(df)[colnames(df) == "Bioreacteur.1"] <- "Jour"
colnames(df)[colnames(df) == "DIC.mg.L."] <- "Moy"
colnames(df)[colnames(df) == "DIC.mg.L..1"] <- "Sd"

head(df, 1)


#### Now, we have to put the number of the days. Let's just use the long (but simple) way using our old friend "c()". 
#### Also, we should look at what we did.
df$Jour <- as.numeric(c("6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24"))

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", " ")))

head(df,15)

#### Alright, everything looks good ^.^ Let's start doing some tests
```

### Visualize the data 
```{r message=TRUE, echo=TRUE}
Box_Dic <-ggplot(df, aes(x = Bioreacteur, y = Moy)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average DIC (mg/L)") +
      theme(text = element_text(size = 14))
Box_Dic
ggsave("C:/Users/patel/OneDrive/Documents/Maitrise/1.Test_1/2.Figures/Box_DIC.jpeg", plot = last_plot())
```

### Let's see if our data follows a normal distribution 
### Let's see if the variation is homogenous 
```{r echo=TRUE}
shapiro.test(df$Moy)
# W = 0.76882, p-value = 1.862e-05
# It does not have a normal distribution. We have to do a Flinger test

fligner.test(Moy ~ Bioreacteur, data = df)
#Fligner-Killeen:med chi-squared = 0.46147, df = 2, p-value = 0.7939
# The variance is homogenous.
```


### Doing the Kruskal-Wallis test 
```{r echo=TRUE}
kruskal.test(Moy ~ Bioreacteur, data = df)

# Kruskal-Wallis chi-squared = 0.54799, df = 2, p-value = 0.7603
# Since p-value is bigger than 0.05, there are no statistical difference between the bioreactors.
```





#_________________________________________|DOC|_________________________________________#

### Create a dataframe with the mean and sd of DOC based on Bioreactor and the day (Jour) 
```{r}
df <- data.frame()

for(i in unique(Analyse$Jour)) {
  
  subset(Analyse, Jour  == i) -> data_Jour
  
  aggregate(DOC.mg.L. ~ Bioreacteur, data = data_Jour, mean, na.rm = TRUE) -> DOC_moy
  
  aggregate(DOC.mg.L. ~ Bioreacteur, data = data_Jour, sd, na.rm = TRUE) ->DOC_sd
  
  df <- rbind(df,data.frame(DOC_moy,DOC_sd))
  
}
#print(df)


#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order + Remove the Tourb and Puit row.
#### We should also have to make sure that the data got rearranged and that the Tourb and Puit row got deleted.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))
df <- df[-c(31,32),]
#tail(df, 20)


#### Let's rename the Bioreacteur.1 column as "Jour" since that column is pretty useless for us.
#### While we're at it, we can also change the name of our other column so that they represent the values properly.
#### Plus, let's also make sure our changes worked
colnames(df)[colnames(df) == "Bioreacteur.1"] <- "Jour"
colnames(df)[colnames(df) == "DOC.mg.L."] <- "Moy"
colnames(df)[colnames(df) == "DOC.mg.L..1"] <- "Sd"

#head(df, 1)

#### Now, we have to put the number of the days. Let's just use the long (but simple) way using our old friend "c()". 
#### Also, we should look at what we did.
df$Jour <- as.numeric(c("6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24"))

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", " ")))
```


### Visualisze the data
```{r}
Box_Doc <-ggplot(df, aes(x = Bioreacteur, y = Moy)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average DOC (mg/L)") +
      theme(text = element_text(size = 14))
Box_Doc
ggsave("C:/Users/patel/OneDrive/Documents/Maitrise/1.Test_1/2.Figures/Box_DOC.jpeg", plot = last_plot())
```


### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous 
```{r}
shapiro.test(df$Moy)
# W = 0.43736, p-value = 1.238e-09

fligner.test(Moy ~ Bioreacteur, data = df)
# Fligner-Killeen:med chi-squared = 0.44481, df = 2, p-value = 0.8006
```


### On va faire le test non-parametrique Kruskal-Wallis 
```{r}
kruskal.test(Moy ~ Bioreacteur, data = df)

# Kruskal-Wallis chi-squared = 3.4245, df = 2, p-value = 0.1805
```





#_________________________________________|NO2|_________________________________________#

### Making the dataframe with the mean and sd
```{r}
df <- data.frame()

for(i in unique(Analyse$Jour)) {
  
  subset(Analyse, Jour  == i) -> data_Jour
  
  aggregate(NNO2.mg.L. ~ Bioreacteur, data = data_Jour, mean, na.rm = TRUE) -> NO2_moy
  
  aggregate(NNO2.mg.L. ~ Bioreacteur, data = data_Jour, sd, na.rm = TRUE) -> NO2_sd
  
  df <- rbind(df,data.frame(NO2_moy,NO2_sd))
  
}

#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order + Remove the Tourb and Puit row.
#### We should also have to make sure that the data got rearranged and that the Tourb and Puit row got deleted.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))
df <- df[-c(31,32),]

#### Let's rename the Bioreacteur.1 column as "Jour" since that column is pretty useless for us.
#### While we're at it, we can also change the name of our other column so that they represent the values properly.
#### Plus, let's also make sure our changes worked
colnames(df)[colnames(df) == "Bioreacteur.1"] <- "Jour"
colnames(df)[colnames(df) == "NNO2.mg.L."] <- "Moy"
colnames(df)[colnames(df) == "NNO2.mg.L..1"] <- "Sd"

#### Now, we have to put the number of the days. Let's just use the long (but simple) way using our old friend "c()". 
#### Also, we should look at what we did.
df$Jour <- as.numeric(c("6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24"))

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", " ")))
```

#### Plotting
```{r}
Box_NO2 <-ggplot(df, aes(x = Bioreacteur, y = Moy)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average NO2 (mg/L)") +
      theme(text = element_text(size = 14))
Box_NO2
```



### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous 
```{r}
shapiro.test(df$Moy)
# W = 0.92911, p-value = 0.0465

fligner.test(Moy ~ Bioreacteur, data = df)
# Fligner-Killeen:med chi-squared = 0.36329, df = 2, p-value = 0.8339
```



### Non-parametrique Kruskal-Wallis 
```{r}
kruskal.test(Moy ~ Bioreacteur, data = df)

# Kruskal-Wallis chi-squared = 0.43457, df = 2, p-value = 0.8047
```








#_________________________________________|NO3|_________________________________________#

### Faire un data frame avec les moyennes et sd de NO3 selon le Bioreacteur et le Jour 
```{r}
df <- data.frame()

for(i in unique(Analyse$Jour)) {
  
  subset(Analyse, Jour  == i) -> data_Jour
  
  aggregate(NNO3.mg.L. ~ Bioreacteur, data = data_Jour, mean, na.rm = TRUE) -> NO3_moy
  
  aggregate(NNO3.mg.L. ~ Bioreacteur, data = data_Jour, sd, na.rm = TRUE) -> NO3_sd
  
  df <- rbind(df,data.frame(NO3_moy,NO3_sd))
  
}

#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order + Remove the Tourb and Puit row.
#### We should also have to make sure that the data got rearranged and that the Tourb and Puit row got deleted.
#### I also realized that there's a negatif value so we'll replace that value with 0.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))
df <- df[-c(31,32),]
df[25,2] = 0.00

#### Let's rename the Bioreacteur.1 column as "Jour" since that column is pretty useless for us.
#### While we're at it, we can also change the name of our other column so that they represent the values properly.
#### Plus, let's also make sure our changes worked
colnames(df)[colnames(df) == "Bioreacteur.1"] <- "Jour"
colnames(df)[colnames(df) == "NNO3.mg.L."] <- "Moy"
colnames(df)[colnames(df) == "NNO3.mg.L..1"] <- "Sd"

#### Now, we have to put the number of the days. Let's just use the long (but simple) way using our old friend "c()". 
#### Also, we should look at what we did.
df$Jour <- as.numeric(c("6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24"))

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", " ")))
```


#### On commence d'abord par juste visualiser les données
```{r}
Box_NO3 <-ggplot(df, aes(x = Bioreacteur, y = Moy)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average NO3 (mg/L)") +
      theme(text = element_text(size = 14))
Box_NO3
```




### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous 
```{r}
shapiro.test(df$Moy)
# W = 0.62258, p-value = 1.392e-07

fligner.test(Moy ~ Bioreacteur, data = df)
# Fligner-Killeen:med chi-squared = 3.487, df = 2, p-value = 0.1749
```



### Non-parametrique Kruskal-Wallis test
```{r}
kruskal.test(Moy ~ Bioreacteur, data = df)
# Kruskal-Wallis chi-squared = 2.1625, df = 2, p-value = 0.3392
```





#_________________________________________|NH4|_________________________________________#

### Making the dataframe with the mean and sd 
```{r}
df <- data.frame()

for(i in unique(Analyse$Jour)) {
  
  subset(Analyse, Jour  == i) -> data_Jour
  
  aggregate(NH4_NH3.mg.L.~ Bioreacteur, data = data_Jour, mean, na.rm = TRUE) -> NH3NH4_moy
  
  aggregate(NH4_NH3.mg.L.~ Bioreacteur, data = data_Jour, sd, na.rm = TRUE) -> NH3NH4_sd
  
  df <- rbind(df,data.frame(NH3NH4_moy,NH3NH4_sd))
  
}

#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order + Remove the Tourb and Puit row.
#### We should also have to make sure that the data got rearranged and that the Tourb and Puit row got deleted.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))
df <- df[-c(31,32),]

#### Let's rename the Bioreacteur.1 column as "Jour" since that column is pretty useless for us.
#### While we're at it, we can also change the name of our other column so that they represent the values properly.
#### Plus, let's also make sure our changes worked
colnames(df)[colnames(df) == "Bioreacteur.1"] <- "Jour"
colnames(df)[colnames(df) == "NH4.mg.L."] <- "Moy"
colnames(df)[colnames(df) == "NH4.mg.L..1"] <- "Sd"

#### Now, we have to put the number of the days. Let's just use the long (but simple) way using our old friend "c()". 
#### Also, we should look at what we did.
df$Jour <- as.numeric(c("6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24","6","8","10","12","14","16","18","20","22","24"))

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", " ")))
```

#### Plotting
```{r}
Box_NH4 <-ggplot(df, aes(x = Bioreacteur, y = Moy)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average NH4 (mg/L)") +
      theme(text = element_text(size = 14))
Box_NH4
```


### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous 
```{r}
shapiro.test(df$Moy)
#W = 0.51371, p-value = 7.504e-09

fligner.test(Moy ~ Bioreacteur, data = df)
# Fligner-Killeen:med chi-squared = 0.98381, df = 2, p-value = 0.6115
```


### Non-parametrique Kruskal-Wallis test
```{r}
kruskal.test(Moy ~ Bioreacteur, data = df)

# Kruskal-Wallis chi-squared = 0.7513, df = 2, p-value = 0.6868
```




#______________________________________|Dissolved O2|______________________________________#

### Making the dataframe for O2 and pH
```{r}

####  Garder la colonne d'O2 et de pH dissous seulement
df <- Analyse[c("Bioreacteur", "Jour", "O2_dis", "pH")]
df <- df[complete.cases(Analyse$pH), ]
df <- df[!(df$Bioreacteur == "Puit"),] 
df <- df[!(df$Bioreacteur == "Tourb"),]

#### We'll change the names in the Bioreacteur column to Bioreactor 1, Bioreactor 2, Bioreactor 3 and Field
df$Bioreacteur <- ifelse(grepl("^B", df$Bioreacteur), "Bioreactor 1",
                        ifelse(grepl("^J", df$Bioreacteur), "Bioreactor 2",
                         ifelse(grepl("^R", df$Bioreacteur), "Bioreactor 3", ""))) 

#### Rearrange the rows to have all "Bleu", "Jaune", "Rouge" in that order.
df <- df %>% arrange(factor(Bioreacteur, levels = c("Bleu", "Jaune", "Rouge")))

```

### Plotting
```{r}
#### Regardons les boxplots
Box_O2 <-ggplot(df, aes(x = Bioreacteur, y = O2_dis)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average % dissolve Oxygen") +
      theme(text = element_text(size = 14))
Box_O2
```

### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous
```{r}
shapiro.test(df$O2_dis)
# W = 0.90024, p-value = 0.01355

fligner.test(O2_dis ~ Bioreacteur, data = df)
# Fligner-Killeen:med chi-squared = 0.38826, df = 2, p-value = 0.8236
```


### Non-parametrique Kruskal-Wallis test
```{r}
kruskal.test(O2_dis ~ Bioreacteur, data = df)
# Kruskal-Wallis chi-squared = 0.41282, df = 2, p-value = 0.8135
```





#_________________________________________|pH|_________________________________________#

### Plotting
```{r}
Box_pH <-ggplot(df, aes(x = Bioreacteur, y = pH)) + 
      geom_boxplot() + 
      xlab("") +
      ylab("Average pH") +
      theme(text = element_text(size = 14))
Box_pH
```

### Let's see if our data follows a normal distribution
### Let's see if the variation is homogenous
```{r}
shapiro.test(df$pH)
# W = 0.96529, p-value = 0.4195

bartlett.test(pH ~ Bioreacteur, data = df)
# Bartlett's K-squared = 2.2439, df = 2, p-value = 0.3256
```

### Non-parametrique Kruskal-Wallis test
```{r}
kruskal.test(pH ~ Bioreacteur, data = df)

# Kruskal-Wallis chi-squared = 5.7204, df = 2, p-value = 0.05726
```


